---
title: "Initial Korean covid 19"
author: "Rita Jarbanda"
date: "Fall 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## Methods

Based on the data we found in KRdata, we will try and find out if there is a difference between men and women. Is there any difference in how covid kills men and women. Is covid more harmful to men? Or is it more harmful to women?
We are going to use the CandP helper files to compare two categorical data types Sex and Results Because this is a cat-cat problem we will first create a bar chart, then we will run favstates to show graphical and numerical results.
For inferential results we are going to run a Chi-Squared test which will tell us if there is a difference in the data but not where. 

## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  When viewing the first chart, it shows that more women got COVID than men.
However when looking at the second chart, the results show that of whole group of people that got COVID, more men were being killed with COVID than women.  

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
``` 

The data above confirmed the data we see in the bar chart ggplot results. Which was that women got COVID at a higher rate than men. However, of people who got covid, more men died than women.  

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

After running our inferential test, we found our p-value to be 0.0009 which is far lower than 0.05.Based on the result of chisqure and fisher test, the result show that if you are a male, you are twice as likely to die from COVID than women.  

## Conclusions

First. we looked at the data available in kRdata and confirmed that this was a category~category problem. After we determined this we knew we need to use or Cat~Cat helper files to run a bar chart using ggplot. After seeing the result of the ggplot, we determined that of the people who get COVID, women appear to get COVID at higher rate. It also appears that of the people who got COVID, men died at a higher rate than women. However, we needed to run more test to confirm.  
